AbxD01 Analysis

### 2015-02-17 abxD01_analysis.Rmd
add SE to bar plots, color by phylum
Update .Rmd to reflect changes in bar chart code

## Figure 1: Bargraph

@@

This figure was built using the R script called "abxD01.barcharts.xOTU.test.r" in the abxD01 folder.The input files are listed in Github. The graph can be sorted 2 ways either by the untreated's decreasing relabund, or by the most abundant within each phylum. For each treatment, the resulting *C. difficile* CFU/g Feces is also labeled on the Y axis below the treatment name. 
The parameters I changed in the r script were:
sortbyphyl<-TRUE
graphbyphyl<-FALSE
abx<-c("Untreated", "Ciprofloxacin", "Clindamycin", "Vancomycin", "Streptomycin", "Cefoperazone", "Ampicillin", "Metronidazole")

File saved as "topdose2_tx1_sorted_10x20.pdf" and "topdose2_tx2_sorted_10x20.pdf"."

@@

##Paper Figure: correlations for topdose 
otuCorrs<-read.csv("~/Documents/Github/abxD01/Figure 2/corr_avgAbove0.001_topdose2_stripchart.csv", header=T)
#labels=c("Lachnospiraceae", "Ruminococcaceae", "Clostridia", "Lactobacillales", "Firmicutes", "Bacillales", "Porphyromonadaceae", "Bacteroidales", "Bacteroidetes", "Actinobacteria", "Proteobacteria", "Anaeroplasma", "Deinococcus", "Unclassified")
#ns=c("n=18", "n=9", "n=8", "n=4", "n=2", "n=1", "n=15", "n=4", "n=1", "n=4", "n=3", "n=1", "n=1", "n=3")

#test<-otuCorrs[otuCorrs$phyl=="Bacteroidetes",]
#length(unique(test$family))

pch <- c(1,1,1,1, 6,6, 6, 6, 6, 6, 6, 6,0, 0, 9, 2, 2, 4)
par(mar=c(12, 7, 2, 2) +0.1, mgp=c(4, 1, 0)) #default is 5.1 4.1 4.1 2.1 [bottom, left, top, right space], mgp=c(3, 1, 0) [label line location for x/y location labels, tick mark labels location, tick mark locations]
stripchart(otuCorrs$corr ~ otuCorrs$graphOrder, vertical=TRUE, ylab="Spearman Correlation", ylim=c(-1, 1), cex.axis=1.1, cex=2, pch=pch, lwd=2, col="black",  method="jitter",  jitter=0.25, cex.lab=1.3, yaxt="n", xaxt="n")
axis(1, cex.axis=1, at=(1:nlevels(otuCorrs$family)), labels=FALSE)
text(1:nlevels(otuCorrs$family)+.1, par("usr")[3]-0.05, label=unique(otuCorrs$family), xpd=NA, pos=2, srt=45, cex=1.2)
#text(1:nlevels(otuCorrs$family), par("usr")[3]-0.01, label=ns, xpd=NA, pos=1, cex=.9)
axis(2, cex.axis=1.1, at=seq(-1,+1, by=.1), las=1)
abline(v=c(1:nlevels(otuCorrs$family)), col="dark gray")
#abline(h=c(.3, -.3), col="dark gray", lty="dashed", lwd=4) #change to something else
abline(h=0, col="black", lwd=2)
abline(h=c(0.25, 0.5, 0.75, -0.25, -0.5, -0.75), col="light gray", lwd=2, lty="dashed")
legend("top", legend=unique(otuCorrs$phyl), pch=c(1, 6, 0, 9, 2, 4), cex=0.7, bty="o", horiz=TRUE)

@@

Figure saved as "corr_avgAbove0.001_topdose2_10x8.pdf".

@@

The code for this is the script "abxD01.barcharts.xOTU.test.r" in the paper's home Github directory. Opening this file, the parameters I changed in the top were:
sortbyphyl<-TRUE
graphbyphyl<-TRUE
abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
file
fileIDS

>>>>

## Figure 1

@@

### Figure 1: Bargraph
This figure was built using the R script called "stacked.barcharts.r", which contains the stackedbarcharts() funciton in the abxD01 folder.The input files are listed in Github. The graph can be sorted 2 ways either by the untreated's decreasing relabund (sortbyphyl=FALSE), or by the most abundant within each phylum (sortbyphyl=TRUE). For each treatment, the resulting *C. difficile* CFU/g Feces is also labeled on the Y axis below the treatment name. 
The parameters I changed in the r script were:
```{r, eval=TRUE}
abx<-c("Untreated", "Ciprofloxacin", "Clindamycin", "Vancomycin", "Streptomycin", "Cefoperazone", "Ampicillin", "Metronidazole")

# Generate the 
stackedbarcharts(file="~/Documents/Github/abxD01/Figure 1/abxD01.final.tx.2.subsample.relabund.topdose2.forlogscale.csv", fileIDS = "~/Documents/Github/abxD01/Figure 1/topdose_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = FALSE)

stackedbarcharts(file="~/Documents/Github/abxD01/Figure 1/abxD01.final.tx.1.subsample.relabund.topdose2.forlogscale.csv", fileIDS = "~/Documents/Github/abxD01/Figure 1/topdose_tx1_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = FALSE)

Files saved as "topdose2_tx1_sorted_10x20.pdf" and "topdose2_tx2_sorted_10x20.pdf".

@@

##Paper Figure: correlations for topdose 
otuCorrs<-read.csv("~/Documents/Github/abxD01/Figure 2/corr_avgAbove0.001_topdose2_stripchart.csv", header=T)

pch <- c(6,6,6,6, 1,1, 1, 1, 1, 1, 1, 1,0, 0, 9, 2, 2, 4)
par(mar=c(12, 7, 2, 2) +0.1, mgp=c(4, 1, 0)) #default is 5.1 4.1 4.1 2.1 [bottom, left, top, right space], mgp=c(3, 1, 0) [label line location for x/y location labels, tick mark labels location, tick mark locations]
stripchart(otuCorrs$corr ~ otuCorrs$graphOrder, vertical=TRUE, ylab="Spearman Correlation", ylim=c(-1, 1), cex.axis=1, cex=2, pch=pch, lwd=2, col="black",  method="jitter",  jitter=.2, cex.lab=1.1, yaxt="n", xaxt="n")
axis(1, cex.axis=1, at=(1:nlevels(otuCorrs$family)), labels=FALSE)
family.labels <- unique(otuCorrs[order(otuCorrs$graphOrder), "family"])
text(1:nlevels(otuCorrs$family)+.1, par("usr")[3]-0.05, label=(family.labels), xpd=NA, pos=2, srt=45, cex=1)
axis(2, cex.axis=1, at=seq(-1,+1, by=.1), las=1)
color_transparent <- adjustcolor("black", alpha.f = 0.3) 
abline(v=c(1:nlevels(otuCorrs$family)), col=color_transparent)
abline(h=0, col="dark gray", lwd=2)
abline(h=c(0.25, 0.5, 0.75, -0.25, -0.5, -0.75), col=color_transparent, lwd=2, lty="dashed")
phylum.labels <- unique(otuCorrs[order(otuCorrs$graphOrder), "phyl"])
leg<-legend("top", legend=phylum.labels, pch=c(6, 1, 0, 9, 2, 4), cex=0.7, bty="o", horiz=TRUE, pt.cex=1.5, text.width=2.5)

@@

Figure saved as "corr_avgAbove0.001_topdose2_10x10.pdf".

@@

The code for this is the script "stacked.barcharts.r" which contains the stackedbarcharts() function in the paper's home Github directory. Opening this file, the parameters I changed in the top were:
```{r, eval=FALSE}
setwd("/Users/alyxschubert/Documents/Github/abxD01/")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv", fileIDS = "./Figure 3/allceftitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrCef")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv", fileIDS = "./Figure 3/allstreptitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrStrep")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv", fileIDS = "./Figure 3/allvanctitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrVanc")
